New lower bounds for Schur and weak Schur numbers

### 2021-04-09 article/article.tex
corrections de légers problèmes

This paper offers new lower bounds for both Schur and weak Schur numbers. These results were produced by pushing forward
Rowley's "templates"-based approach for Ramsey and Schur numbers in 2020. Finding suitable templates allows us to apply Rowley's construction
to get explicit partitions improving lower bounds. Furthermore, this paper tries to analyse former work on the subject based on
the principle that good partitions into \(n+1\) subsets start with good partitions into \(n\) subsets. We show that exceeding
the previous lower bound \(WS(6) \geqslant 582\) is impossible with such a method upon imposing certain conditions on the good
5-subsets partition. The new lower bounds includes \(S(9) \geqslant 17803\), \(S(10) \geqslant 60948\), \(WS(9) \geqslant 22536 \) and \(WS(10) \geqslant 71214 \).

@@

\par
These new results come from Rowley's approach for Ramsey graphs which is relatively new. Therefore, it would not surprise us
if lower bounds are later improved using better templates. Moreover, studying specifically \(S_+(n)\) and \(WS_+(n)\) numbers might be
of interest as they are closely related to Schur and weak Schur numbers.

@@

\par
We would first like to address warm thanks to Fred Rowley who kindly answered our many questions and with who it was a pleasure
to exchange. A great thank-you to Marijn Heule who also took the time to respond to our requests. Finally, last but not least, we would
like to thank our professors, Joanna Tomasik and Arpad Rimmel, who guided and advised us during the whole research and writing.

>>>>

This paper offers new lower bounds for both Schur and weak Schur numbers. These results were produced by pushing forward
Rowley's "templates"-based approach for Ramsey and Schur numbers in 2020. Finding suitable templates allows us to apply Rowley's construction
to get explicit partitions improving lower bounds. We also developed a new method to improve weak Schur lower bounds. Furthermore, this paper tries to analyse former work on the subject based on
the principle that good partitions into \(n+1\) subsets start with good partitions into \(n\) subsets. We show that exceeding
the previous lower bound \(WS(6) \geqslant 582\) is impossible with such an assumption upon imposing certain conditions on the good
5-subsets partition. The new lower bounds includes \(S(9) \geqslant 17803\), \(S(10) \geqslant 60948\), \(WS(9) \geqslant 22536 \) and \(WS(10) \geqslant 71214 \).

@@

\par
These new results come from Rowley's approach for Ramsey graphs which is relatively new. Therefore, it would not surprise us
if lower bounds are later improved using better templates. Moreover, studying specifically \(S^+(n)\) and \(WS^+(n)\) numbers might be
of interest as they are closely related to Schur and weak Schur numbers.

@@

\par
We would first like to thank our professors, Joanna Tomasik and Arpad Rimmel, who guided and advised 
us during the whole research and writing. Finally, last but not least, we would like to address warm thanks 
to Fred Rowley who kindly answered our many questions and with who it was a pleasure to exchange.

### 2021-04-09 article/weakschur.tex
tableau+modif pi
corrections de légers problèmes

\renewcommand{\arraystretch}{2}
\begin{tabular}{|c|c|c|c|c|c|c|c|c|>{\columncolor{blue}}c|>{\columncolor{blue}}c|>{\columncolor{blue}}c|>{\columncolor{green}} c|}
\hline
  & & & & & \cellcolor{green}1 & \cellcolor{green}2 & \cellcolor{blue} ...&\cellcolor{green}\(r\)&\(r+1\)&...&\(b - 1\)&\(b\) \\
\hline
   \cellcolor{red}9 & \cellcolor{red}10 & \cellcolor{red}11 & \cellcolor{red}12 &\cellcolor{red}13&\cellcolor{red}14&\cellcolor{red}15&\cellcolor{red}16&\cellcolor{red}17&18&19&20&21\\
\hline
\cellcolor{yellow}22&\cellcolor{yellow}23&\cellcolor{yellow}24&\cellcolor{yellow}25&\cellcolor{yellow}26&\cellcolor{yellow}27&\cellcolor{yellow}28&\cellcolor{yellow}29&\cellcolor{yellow}30&31&32&33&34\\
\hline
...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...&...&...&...\\
\hline
\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&...&...&...&...\\
\hline
...&...&...&...&...&...&...&...&...&...&...&...&...\\
\hline
\end{tabular}

@@

We will now define notations and results, we will use in the following theorem.

\begin{notation}

Let (a,b) \(n \in (\mathbb{N}^*)^2\),\(a>b\), we will define \(\pi_{a,b}\) the projection:
\[ \pi_{a,b}:x->(Id+a\mathbf{1}_{ [\![0,b]\!]})(\text{x mod a}))\]
\end{notation}

@@

\begin{notation}

Let (a,b) \(n \in (\mathbb{N}^*)^2\),\(a>b\), we will define \(\lambda_{a,b}\) the projection:
\[ \lambda_{a,b}:x->1+ \left\lfloor\dfrac{x-b-1}{a}\right\rfloor\]
\end{notation}

@@

\begin{remark}
In the following theorem, \(\lambda\) is the function which return the line number of an element x.
\end{remark}

@@

if x mod a \(\leqslant b\):
\(a\lambda(x)+\pi(x)-a=a(\left\lfloor\dfrac{x}{a}\right\rfloor-1)+\text{x mod a}+a=x\)
\begin{definition}
Let \( (p,k,b) \in (\mathbb{N}^*)^3\), Let \((A_1,....,A_k)\) a partition of  \([\![1, p]\!]\).
This partition is said to be a b-weakly-sum-free template (b-WSF-template) of \(k\) colors and lenght \(p\) when:
\\\\
\underline{\(\forall i \in [\![1, k]\!], \quad A_i\) is weakly-sum-free}
\\\\
\underline{\(\forall i \in [\![1, k]\!], \quad A_i\backslash [\![1, b]\!]\) is sum-free}
\\\\
\underline{For \(A_k\) (the special subset):} \quad \(\forall (x,y) \in A_k^2,\)
\\
\[if \quad x+y>b+2(p-b),\quad x+y-2(p-b)\notin A_k\]
\\
\underline{For the others subsets:}\quad \(\forall i \in [\![0,k-1]\!], \forall(x,y) \in A_i^2\)
\\
\[
(Id+(p-b)\mathbf{1}_{ [\![0,b]\!]})(x+y ~\text{mod}~ (p-b)) \notin A_i
\]
\end{definition}

@@

\renewcommand{\arraystretch}{2}
\begin{center}
\[
\begin{array}{|c|c|c|c|c|c|c|c|c|>{\columncolor{blue}}c|>{\columncolor{blue}}c|>{\columncolor{blue}}c|>{\columncolor{green}} c|}
\cline{6-9}
   \multicolumn{5}{c|}{} & \cellcolor{green}1 & \cellcolor{green}2 & \cellcolor{blue} 3&\cellcolor{green}4&5&6&7&8 \\
\hline
   \cellcolor{red}9 & \cellcolor{red}10 & \cellcolor{red}11 & \cellcolor{red}12 &\cellcolor{red}13&\cellcolor{red}14&\cellcolor{red}15&\cellcolor{red}16&\cellcolor{red}17&18&19&20&21\\
\hline
\cellcolor{yellow}22&\cellcolor{yellow}23&\cellcolor{yellow}24&\cellcolor{yellow}25&\cellcolor{yellow}26&\cellcolor{yellow}27&\cellcolor{yellow}28&\cellcolor{yellow}29&\cellcolor{yellow}30&31&32&33&34\\
\hline
...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...&...&...&...\\
\hline
\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&...&...&...&...\\
\hline
...&...&...&...&...&...&...&...&...&...&...&...&...\\
\hline
\end{array}
\]
\end{center}
\textsc{Proof :}Let \((q,n,b) \in (\mathbb{N}^*)^3\), let \( (p,k) \in (\mathbb{N}^*)^2\), let a=q-b. \\
We denote by \(g\) the projection of the equivalence relation induced by the partition of \([\![1,q]\!]\) and \(h\) the
one induced by the partition of \([\![1,p]\!]\). Each equivalence class is represented by a single integer, therefore :
\[ g : [\![1,q]\!] \longrightarrow [\![1,n]\!] \text{ and } (A_{g^{-1}(1)},...A_{g^{-1}(q)})\text{ is a b-WSF-template.}
\]
\[h : [\![1,p]\!] \longrightarrow [\![1,k]\!] \text{ and } \forall (x,y) \in [\![1,q]\!]^2 \text{, } h(x) = h(y)
\Longrightarrow h(x+y) \neq h(x)
\]

>>>>

\renewcommand{\arraystretch}{2}
\begin{center}
\[
\begin{array}{|c|c|c|c|c|c|c|c|c|>{\columncolor{blue}}c|>{\columncolor{blue}}c|>{\columncolor{blue}}c|>{\columncolor{green}} c|}
\cline{6-9}
   \multicolumn{5}{c|}{} & \cellcolor{green}1 & \cellcolor{green}2 & \cellcolor{blue} 3&\cellcolor{green}4&5&6&7&8 \\
\hline
   \cellcolor{red}9 & \cellcolor{red}10 & \cellcolor{red}11 & \cellcolor{red}12 &\cellcolor{red}13&\cellcolor{red}14&\cellcolor{red}15&\cellcolor{red}16&\cellcolor{red}17&18&19&20&21\\
\hline
\cellcolor{yellow}22&\cellcolor{yellow}23&\cellcolor{yellow}24&\cellcolor{yellow}25&\cellcolor{yellow}26&\cellcolor{yellow}27&\cellcolor{yellow}28&\cellcolor{yellow}29&\cellcolor{yellow}30&31&32&33&34\\
\hline
...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...&...&...&...\\
\hline
\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&...&...&...&...\\
\hline
...&...&...&...&...&...&...&...&...&...&...&...&...\\
\hline
\end{array}
\]
\end{center}

@@

We will now define functions and results, that will be used in the following theorem.

\begin{definition}

Let (a,b) \(n \in (\mathbb{N}^*)^2\),\(a>b\), we will define \(\pi_{a,b}\) the projection:
\[ \pi_{a,b}:x \longmapsto (Id+a\mathbf{1}_{ [\![0,b]\!]})(\text{x mod a}))\]
\end{definition}

@@

\begin{definition}

Let (a,b) \(n \in (\mathbb{N}^*)^2\),\(a>b\), we will define \(\lambda_{a,b}\) the projection:
\[ \lambda_{a,b}:x \longmapsto 1+ \left\lfloor\dfrac{x-b-1}{a}\right\rfloor\]
\end{definition}

@@

\begin{remark}
In the following theorem, \(\lambda\) is the function which maps an element x to its line number.
\end{remark}

@@

if x mod a \(\leqslant b\):
\(a\lambda(x)+\pi(x)-a=a(\left\lfloor\dfrac{x}{a}\right\rfloor-1)+\text{x mod a}+a=x\)

\begin{definition}
Let \( (p,k,b) \in (\mathbb{N}^*)^3\), Let \((A_1,....,A_k)\) a partition of  \([\![1, p]\!]\).
This partition is said to be a b-weakly-sum-free template (b-WSF-template) of \(k\) colors and lenght \(p\) when:
\\\\
\underline{\(\forall i \in [\![1, k]\!], \quad A_i\) is weakly-sum-free}
\\\\
\underline{\(\forall i \in [\![1, k]\!], \quad A_i\backslash [\![1, b]\!]\) is sum-free}
\\\\
\underline{For \(A_k\) (the special subset):} \quad \(\forall (x,y) \in A_k^2,\)
\\
\[if \quad x+y>b+2(p-b),\quad x+y-2(p-b)\notin A_k\]
\\
\underline{For the others subsets:}\quad \(\forall i \in [\![0,k-1]\!], \forall(x,y) \in A_i^2\)
\\
\[
x+y>p \implies \pi(x+y) \notin A_i
\]
\end{definition}

@@

\textsc{Proof :}Let \((q,n,b) \in (\mathbb{N}^*)^3\), let \( (p,k) \in (\mathbb{N}^*)^2\), let a=q-b. \\
We denote by \(g\) the projection of the equivalence relation induced by the partition of \([\![1,q]\!]\) and \(h\) the
one induced by the partition of \([\![1,p]\!]\). Each equivalence class is represented by a single integer, therefore :
\[ g : [\![1,q]\!] \longrightarrow [\![1,n]\!] \text{ and } (A_{g^{-1}(1)},...A_{g^{-1}(q)})\text{ is a b-WSF-template.}
\]
\[h : [\![1,p]\!] \longrightarrow [\![1,k]\!] \text{ and } \forall (x,y) \in [\![1,q]\!]^2 \text{, } h(x) = h(y)
\Longrightarrow h(x+y) \neq h(x)
\]
